---
title: "HELPrct (Data Verbs) Activity"
author: "Hanbyeol Lee"
date: "May 30"
output: html_notebook
---

## Set Up:

```{r}
# load packages
library(tidyverse)  # for data verbs like `group-by()` & `summarise()`
library(mosaicData) # includes `HELPrct` data set 
library(mosaic)     # includes `mplot()` interactive plot builder
library(esquisse)   # includes `esquisser()` interactive plot builder


# Load the HELPrct data from `mosaicData` into our RStudio environment
data("HELPrct", package = "mosaicData")

```

## Data Description

*TASK 1: Write a couple of sentences describing the `HELPrct` data--What is the setting for the data? What are they about?* I got the description from help function: These are survey data collected by the US National Center for Health Statistics (NCHS). Each year, approximately 5,000 individuals are interviewed in their homes. The data cover 1999-2004.

Note that the NHANES package provides data from 2009-2012, including more health, life-style, and socio-economic variables.

The Format has 31,126 observations on the following 31 variables, sex, age, pregnant, ethnicity, death, floowup, smoker, diabetic, etc.

## Basic Summaries with `summarise()`

*TASK 2: Produce an expression involving `summarize()` and `HELPrct` that will produce the following.*

1.  *number of people (cases) in `HELPrct` study*
2.  *combined total number of times in the past 6 months all the people in `HELPrct` entered a detox program (measured at baseline)*
3.  *mean time (in days) to first use of any substance post-detox for all the people in `HELPrct`*

```{r}

summary_stats <- HELPrct %>%
  summarise(
    num_people = n(),
    total_detox_entries = sum(e2b, na.rm = TRUE),
    mean_time_to_first_use = mean(daysanysub, na.rm = TRUE)
  )

head(summary_stats)

```

## Subgrouping with `group_by()`

*TASK 3: Repeat Task 2 above, but calculate the results group-by-group. Show your code and write a sentence or two about what you observe in the results for each of the following:*

1.  *males versus females*

```{r}
groupby_summary_stats <- HELPrct %>%
  group_by(sex)  %>%
  summarise(
   num_people = n(),
    total_detox_entries = sum(e2b, na.rm = TRUE),
    mean_time_to_first_use = mean(daysanysub, na.rm = TRUE)
  )

head(groupby_summary_stats)

```

2.  *homeless or not*

```{r}
groupby_summary_stats <- HELPrct %>%
  group_by(homeless)  %>%
  summarise(
   num_people = n(),
    total_detox_entries = sum(e2b, na.rm = TRUE),
    mean_time_to_first_use = mean(daysanysub, na.rm = TRUE)
  )

head(groupby_summary_stats)
```

3.  *substance*

```{r}
groupby_summary_stats <- HELPrct %>%
  group_by(substance)  %>%
  summarise(
   num_people = n(),
    total_detox_entries = sum(e2b, na.rm = TRUE),
    mean_time_to_first_use = mean(daysanysub, na.rm = TRUE)
  )

head(groupby_summary_stats)
```

4.  *break down the homeless versus housed further, by sex*

```{r}
groupby_summary_stats <- HELPrct %>%
  group_by(homeless,sex)  %>%
  summarise(
   num_people = n(),
    total_detox_entries = sum(e2b, na.rm = TRUE),
    mean_time_to_first_use = mean(daysanysub, na.rm = TRUE)
  )

head(groupby_summary_stats)
```

5.  *break down the homeless versus housed further, by substance*

```{r}
groupby_summary_stats <- HELPrct %>%
  group_by(homeless,substance)  %>%
  summarise(
   num_people = n(),
    total_detox_entries = sum(e2b, na.rm = TRUE),
    mean_time_to_first_use = mean(daysanysub, na.rm = TRUE)
  )

head(groupby_summary_stats)
```

## TASK 4:

*Include one or more interesting plots of the data involving at least 3 variables per plot. Write a few sentences to explain the story that your plot tells about these data. You can use one of the relationships that you studied in Task 2, or you can explore a different group of variables in the `HELPrct` that show something interesting. You may use interactive tools like `esquisser` or `mplot` but you must copy/paste the `ggplot2` code to create the plot here (i.e., do not embed a screenshot or image file).*

# The boxplot shows the age distribution for each primary substance of abuse.

```{r}
ggplot(HELPrct) +
  aes(x = age, y = substance, colour = sex) +
  geom_boxplot(fill = "#112446") +
  scale_color_hue(direction = 1) +
  theme_minimal()
```
