---
title: "Economic Indicators and Cryptocurrency Prices"
author: "Your Name"
date: "2024-06-16"
output: html_document
---

## Guiding Research Question
How do economic indicators like PMI, Service PMI, and Consumer Confidence influence the prices of Bitcoin and Ethereum?

## Each Data Intake 

```{r}
# Load required libraries
library(tidyverse)

# Load data

bitcoin_historical <- read_csv("~/STAT184/stat184-test/184-test/BTC-USD (2014-2024).csv")
ethereum_historical <- read_csv("~/STAT184/stat184-test/184-test/ETH-USD (2017-2024).csv")
economic_data <- read_csv("~/STAT184/stat184-test/184-test/economic data.csv")
```


## Examine each data source

# BTC Prices Jan to May 2024
```{r}

# Check structure and column names
str(bitcoin_historical)
colnames(bitcoin_historical)

# Convert Date column to Date type if necessary
bitcoin_historical$Date <- as.Date(bitcoin_historical$Date, format = "%Y-%m-%d")

# Convert relevant columns to numeric and round to nearest hundredth
bitcoin_historical <- bitcoin_historical %>%
  mutate(across(c(Open, High, Low, Close, `Adj Close`), ~ round(as.numeric(.), 2)))

# Summary and plots
summary(bitcoin_historical)
```
```{r}
# Plot Bitcoin Historical Prices (2014-2024)
ggplot(bitcoin_historical, aes(x = Date, y = Close)) + 
  geom_line() + 
  ggtitle("Bitcoin Prices (2014-2024)") +
  xlab("Date") + 
  ylab("Close Price")
```

```{r}
#potential outliers

boxplot(bitcoin_historical$Close, main = "Boxplot of Bitcoin Closing Prices (2014-2024)", ylab = "Close Price")

```
<Origin and Purpose>

Who: This data is provided by a cryptocurrency exchange or financial data provider.
What: It includes historical Bitcoin prices from 2014 to 2024.
When: The data covers daily Bitcoin prices from January 1, 2014, to May 31, 2024.
Where: This data is sourced from online cryptocurrency exchanges or financial data services.
Why: The purpose of this data is to analyze long-term price trends of Bitcoin.
How: The data was collected from the historical records of Bitcoin transactions on various exchanges.

<Key Variables>

Date: Date of the recorded price (Type: Date)
Open: Opening price of Bitcoin on the given date (Type: Numeric)
High: Highest price of Bitcoin on the given date (Type: Numeric)
Low: Lowest price of Bitcoin on the given date (Type: Numeric)
Close: Closing price of Bitcoin on the given date (Type: Numeric)
Adj Close: Adjusted closing price of Bitcoin, accounting for splits and dividends (Type: Numeric)
Volume: Number of Bitcoin traded on the given date (Type: Numeric)

<Potential Outliers>
The price ranges around 60k were the Outliers which they were All Time High prices were in 2021. 

```



# ETH Prices 2017-2024
```{r}
# Check structure and column names
str(ethereum_historical)
colnames(ethereum_historical)

# Convert Date column to Date type if necessary
ethereum_historical$Date <- as.Date(ethereum_historical$Date, format = "%Y-%m-%d")

# Convert relevant columns to numeric and round to nearest hundredth
ethereum_historical <- ethereum_historical %>%
  mutate(across(c(Open, High, Low, Close, `Adj Close`), ~ round(as.numeric(.), 2)))

# Summary and plots
summary(ethereum_historical)

```
```{r}
# Plot Ethereum Historical Prices (2017-2024)
ggplot(ethereum_historical, aes(x = Date, y = Close)) + 
  geom_line() + 
  ggtitle("Ethereum Prices (2017-2024)") +
  xlab("Date") + 
  ylab("Close Price")

```
# Economic Indicators Over Time
```{r}
# Check structure and column names
str(economic_data)
colnames(economic_data)

# Convert Date column to Date type if necessary
economic_data$date <- as.Date(economic_data$date, format = "%Y-%m-%d")

# Convert relevant columns to numeric and round to nearest hundredth
economic_data <- economic_data %>%
  mutate(across(c(`manufacturing pmi`, `services pmi`, `consumer confidence`), ~ round(as.numeric(.), 2)))

# Summary and plots
summary(economic_data)

# Plot Economic Indicators Over Time
ggplot(economic_data, aes(x = date)) +
  geom_line(aes(y = `manufacturing pmi`, color = "Manufacturing PMI")) +
  geom_line(aes(y = `services pmi`, color = "Services PMI")) +
  geom_line(aes(y = `consumer confidence`, color = "Consumer Confidence")) +
  ggtitle("Economic Indicators Over Time") +
  xlab("Date") +
  ylab("Indicator Value") +
  scale_color_manual(values = c("Manufacturing PMI" = "blue", "Services PMI" = "green", "Consumer Confidence" = "red"))

```
##Comprarison BTC for each economic Indicator to see the trend
```{r}
# Filter economic data to include only dates from 2020 to 2024
economic_data_filtered <- economic_data %>%
  filter(date >= as.Date("2020-01-01") & date <= as.Date("2024-12-31"))

# Filter Bitcoin historical data to include only dates from 2020 to 2024
bitcoin_historical_filtered <- bitcoin_historical %>%
  filter(Date >= as.Date("2020-01-01") & Date <= as.Date("2024-12-31"))

# Merge the filtered data
bitcoin_data <- merge(bitcoin_historical_filtered, economic_data_filtered, by.x = "Date", by.y = "date")

# Rescale Manufacturing PMI to match the range of Bitcoin Prices
max_price <- max(bitcoin_data$Close, na.rm = TRUE)
min_price <- min(bitcoin_data$Close, na.rm = TRUE)

bitcoin_data <- bitcoin_data %>%
  mutate(manufacturing_pmi_scaled = scales::rescale(`manufacturing pmi`, to = c(min_price, max_price)))

# Plot with rescaled Manufacturing PMI
ggplot(bitcoin_data, aes(x = Date)) +
  geom_line(aes(y = Close, color = "Bitcoin Price")) +
  geom_line(aes(y = manufacturing_pmi_scaled, color = "Manufacturing PMI (scaled)")) +
  ggtitle("Bitcoin Price and Manufacturing PMI Over Time (2020-2024)") +
  xlab("Date") +
  ylab("Value") +
  scale_color_manual(values = c("Bitcoin Price" = "black", "Manufacturing PMI (scaled)" = "blue"))


```

```{r}
# Calculate correlation and fit a linear model
lm_manufacturing <- lm(Close ~ `manufacturing pmi`, data = bitcoin_data)
r_squared_manufacturing <- summary(lm_manufacturing)$r.squared

# Plot with rescaled Manufacturing PMI and regression line
ggplot(bitcoin_data, aes(x = `manufacturing pmi`, y = Close)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle(paste("Bitcoin Price vs. Manufacturing PMI\nR-squared =", round(r_squared_manufacturing, 2))) +
  xlab("Manufacturing PMI") +
  ylab("Bitcoin Price")


```
```{r}
# Calculate correlation and fit a linear model
lm_services <- lm(Close ~ `services pmi`, data = bitcoin_data)
r_squared_services <- summary(lm_services)$r.squared

# Plot with rescaled Services PMI and regression line
ggplot(bitcoin_data, aes(x = `services pmi`, y = Close)) +
  geom_point() +
  geom_smooth(method = "lm", color = "green") +
  ggtitle(paste("Bitcoin Price vs. Services PMI\nR-squared =", round(r_squared_services, 2))) +
  xlab("Services PMI") +
  ylab("Bitcoin Price")


```

```{r}
# Calculate correlation and fit a linear model
lm_consumer <- lm(Close ~ `consumer confidence`, data = bitcoin_data)
r_squared_consumer <- summary(lm_consumer)$r.squared

# Plot with rescaled Consumer Confidence and regression line
ggplot(bitcoin_data, aes(x = `consumer confidence`, y = Close)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  ggtitle(paste("Bitcoin Price vs. Consumer Confidence\nR-squared =", round(r_squared_consumer, 2))) +
  xlab("Consumer Confidence") +
  ylab("Bitcoin Price")


```



#Line Plot for Bitcoin Price and Manufacturing PMI

```{r}
# Rescale Manufacturing PMI to match the range of Bitcoin Prices
max_price <- max(bitcoin_data$Close, na.rm = TRUE)
min_price <- min(bitcoin_data$Close, na.rm = TRUE)

bitcoin_data <- bitcoin_data %>%
  mutate(manufacturing_pmi_scaled = scales::rescale(`manufacturing pmi`, to = c(min_price, max_price)))

# Plot with rescaled Manufacturing PMI
ggplot(bitcoin_data, aes(x = Date)) +
  geom_line(aes(y = Close, color = "Bitcoin Price")) +
  geom_line(aes(y = manufacturing_pmi_scaled, color = "Manufacturing PMI (scaled)")) +
  ggtitle("Bitcoin Price and Manufacturing PMI Over Time") +
  xlab("Date") +
  ylab("Value") +
  scale_color_manual(values = c("Bitcoin Price" = "black", "Manufacturing PMI (scaled)" = "blue"))



```
#Line Plot for Bitcoin Price and Service PMI
```{r}
# Rescale Services PMI to match the range of Bitcoin Prices
bitcoin_data <- bitcoin_data %>%
  mutate(services_pmi_scaled = scales::rescale(`services pmi`, to = c(min_price, max_price)))

# Plot with rescaled Services PMI
ggplot(bitcoin_data, aes(x = Date)) +
  geom_line(aes(y = Close, color = "Bitcoin Price")) +
  geom_line(aes(y = services_pmi_scaled, color = "Services PMI (scaled)")) +
  ggtitle("Bitcoin Price and Services PMI Over Time") +
  xlab("Date") +
  ylab("Value") +
  scale_color_manual(values = c("Bitcoin Price" = "black", "Services PMI (scaled)" = "green"))




```
