---
title: "Economic Indicators and BTC Prices"
author: "Hanbyeol Lee"
date: "2024-06-16"
output: html_document
---

## Guiding Research Question

How do economic indicators like PMI, Service PMI, and Consumer Confidence influence the prices of Bitcoin?

## Each Data Intake

```{r}
# Load required libraries
library(tidyverse)

# Load data

bitcoin_historical <- read_csv("~/STAT184/stat184-test/184-test/BTC-USD (2014-2024).csv")
economic_data <- read_csv("~/STAT184/stat184-test/184-test/economic data.csv")
```

## Examine each data source

# \<BTC Prices 2014-2024\>

```{r}

# Check structure and column names
str(bitcoin_historical)
colnames(bitcoin_historical)

# Convert Date column to Date type if necessary
bitcoin_historical$Date <- as.Date(bitcoin_historical$Date, format = "%Y-%m-%d")

# Convert relevant columns to numeric and round to nearest hundredth
bitcoin_historical <- bitcoin_historical %>%
  mutate(across(c(Open, High, Low, Close, `Adj Close`), ~ round(as.numeric(.), 2)))

# Summary statistics
summary(bitcoin_historical)

```

```{r}
# Plot Bitcoin Historical Prices (2014-2024)
ggplot(bitcoin_historical, aes(x = Date, y = Close)) + 
  geom_line() + 
  ggtitle("Bitcoin Prices (2014-2024)") +
  xlab("Date") + 
  ylab("Close Price")

```

```{r}
# Potential outliers
boxplot(bitcoin_historical$Close, main = "Boxplot of Bitcoin Closing Prices (2014-2024)", ylab = "Close Price")
outliers <- boxplot.stats(bitcoin_historical$Close)$out
outliers

```

<Origin and Purpose>

Who: This data is provided by a cryptocurrency exchange or financial data provider. What: It includes historical Bitcoin prices from 2014 to 2024. When: The data covers daily Bitcoin prices from January 1, 2014, to May 31, 2024. Where: This data is sourced from online cryptocurrency exchanges or financial data services. Why: The purpose of this data is to analyze long-term price trends of Bitcoin. How: The data was collected from the historical records of Bitcoin transactions on various exchanges.

<Key Variables>

Date: Date of the recorded price (Type: Date) Open: Opening price of Bitcoin on the given date (Type: Numeric) High: Highest price of Bitcoin on the given date (Type: Numeric) Low: Lowest price of Bitcoin on the given date (Type: Numeric) Close: Closing price of Bitcoin on the given date (Type: Numeric) Adj Close: Adjusted closing price of Bitcoin, accounting for splits and dividends (Type: Numeric) Volume: Number of Bitcoin traded on the given date (Type: Numeric)

<Potential Outliers> The price ranges around 60k were the Outliers which they were All Time High prices were in 2021.

# Economic Indicators Over Time including: origin & purpose, explain key variables, and potential outliers.

```{r}
# Check structure and column names
str(economic_data)
colnames(economic_data)

# Convert date column to Date type if necessary
economic_data$date <- as.Date(economic_data$date, format = "%Y-%m-%d")

# Convert relevant columns to numeric and round to nearest hundredth
economic_data <- economic_data %>%
  mutate(across(c(`manufacturing pmi`, `services pmi`, `consumer confidence`), ~ round(as.numeric(.), 2)))

# Summary statistics
summary(economic_data)

```

```{r}
# Plot Economic Indicators Over Time
ggplot(economic_data, aes(x = date)) +
  geom_line(aes(y = `manufacturing pmi`, color = "Manufacturing PMI")) +
  geom_line(aes(y = `services pmi`, color = "Services PMI")) +
  geom_line(aes(y = `consumer confidence`, color = "Consumer Confidence")) +
  ggtitle("Economic Indicators Over Time") +
  xlab("Date") +
  ylab("Indicator Value") +
  scale_color_manual(values = c("Manufacturing PMI" = "blue", "Services PMI" = "green", "Consumer Confidence" = "red"))


```

```{r}

# Convert relevant columns to numeric + summary
economic_data <- economic_data %>%
  mutate(across(c(`manufacturing pmi`, `services pmi`, `consumer confidence`), as.numeric))

# Summary statistics
summary(economic_data)
```

```{r}
# Boxplot to check for potential outliers in Manufacturing PMI
ggplot(economic_data, aes(x = "", y = `manufacturing pmi`)) +
  geom_boxplot() +
  ggtitle("Boxplot of Manufacturing PMI") +
  ylab("Manufacturing PMI")


```

```{r}
# Identify potential outliers in Manufacturing PMI
outliers_manufacturing <- boxplot.stats(economic_data$`manufacturing pmi`)$out
outliers_manufacturing

```

```{r}
# Boxplot to check for potential outliers in Services PMI
ggplot(economic_data, aes(x = "", y = `services pmi`)) +
  geom_boxplot() +
  ggtitle("Boxplot of Services PMI") +
  ylab("Services PMI")
```

```{r}
# Identify potential outliers in Services PMI
outliers_services <- boxplot.stats(economic_data$`services pmi`)$out
outliers_services
```

```{r}
# Boxplot to check for potential outliers in Consumer Confidence
ggplot(economic_data, aes(x = "", y = `consumer confidence`)) +
  geom_boxplot() +
  ggtitle("Boxplot of Consumer Confidence") +
  ylab("Consumer Confidence")
```

```{r}
# Identify potential outliers in Consumer Confidence
outliers_consumer <- boxplot.stats(economic_data$`consumer confidence`)$out
outliers_consumer
```

<Origin and Purpose>

Who: This data is provided by financial data providers or governmental agencies. What: It includes various economic indicators like PMI, Service PMI, and Consumer Confidence. When: The data covers economic indicators from January 1, 2020, to May 31, 2024. Where: This data is sourced from official reports, surveys, and financial data services. Why: The purpose of this data is to analyze the impact of economic indicators on cryptocurrency prices. How: The data was collected through surveys, reports, and financial data aggregation services.

<Key Variables>

date: Date of the recorded economic indicator (Type: Date) manufacturing pmi: Purchasing Managers' Index for manufacturing (Type: Numeric) services pmi: Purchasing Managers' Index for services (Type: Numeric) consumer confidence: Consumer Confidence Index (Type: Numeric)

<Potential Outliers> There were Service and Manufacturing PMI had several outliers, but not seen in consumer confidence.



# Calculation of R\^2 to see the correlation between each of economic indicator and BTC price. 
```{r}
#Filter economic data to include only dates from 2020 to 2024
economic_data_filtered <- economic_data %>%
  filter(date >= as.Date("2020-01-01") & date <= as.Date("2024-12-31"))

# Filter Bitcoin historical data to include only dates from 2020 to 2024
bitcoin_historical_filtered <- bitcoin_historical %>%
  filter(Date >= as.Date("2020-01-01") & Date <= as.Date("2024-12-31"))

# Merge the filtered data
bitcoin_data <- merge(bitcoin_historical_filtered, economic_data_filtered, by.x = "Date", by.y = "date")
```

```{r}
# Calculate correlation and fit a linear model

lm_manufacturing <- lm(Close ~ `manufacturing pmi`, data = bitcoin_data)
r_squared_manufacturing <- summary(lm_manufacturing)$r.squared

# Plot with regression line
ggplot(bitcoin_data, aes(x = `manufacturing pmi`, y = Close)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle(paste("Bitcoin Price vs. Manufacturing PMI\nR-squared =", round(r_squared_manufacturing, 2))) +
  xlab("Manufacturing PMI") +
  ylab("Bitcoin Price")


```

```{r}
# Calculate correlation and fit a linear model
lm_services <- lm(Close ~ `services pmi`, data = bitcoin_data)
r_squared_services <- summary(lm_services)$r.squared

# Plot with regression line
ggplot(bitcoin_data, aes(x = `services pmi`, y = Close)) +
  geom_point() +
  geom_smooth(method = "lm", color = "green") +
  ggtitle(paste("Bitcoin Price vs. Services PMI\nR-squared =", round(r_squared_services, 2))) +
  xlab("Services PMI") +
  ylab("Bitcoin Price")


```

```{r}
# Calculate correlation and fit a linear model
lm_consumer <- lm(Close ~ `consumer confidence`, data = bitcoin_data)
r_squared_consumer <- summary(lm_consumer)$r.squared

# Plot with rescaled Consumer Confidence and regression line
ggplot(bitcoin_data, aes(x = `consumer confidence`, y = Close)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  ggtitle(paste("Bitcoin Price vs. Consumer Confidence\nR-squared =", round(r_squared_consumer, 2))) +
  xlab("Consumer Confidence") +
  ylab("Bitcoin Price")


```

# We will make a comparison chart: Manufacturing PMI vs BTC Price since it has the highest correlated percentage.

```{r}
# Filter economic data to include only dates from 2020 to 2024
economic_data_filtered <- economic_data %>%
  filter(date >= as.Date("2020-01-01") & date <= as.Date("2024-12-31"))

# Filter Bitcoin historical data to include only dates from 2020 to 2024
bitcoin_historical_filtered <- bitcoin_historical %>%
  filter(Date >= as.Date("2020-01-01") & Date <= as.Date("2024-12-31"))

# Merge the filtered data
bitcoin_data <- merge(bitcoin_historical_filtered, economic_data_filtered, by.x = "Date", by.y = "date")

# Rescale Manufacturing PMI to match the range of Bitcoin Prices
max_price <- max(bitcoin_data$Close, na.rm = TRUE)
min_price <- min(bitcoin_data$Close, na.rm = TRUE)

bitcoin_data <- bitcoin_data %>%
  mutate(manufacturing_pmi_scaled = scales::rescale(`manufacturing pmi`, to = c(min_price, max_price)))

# Plot with rescaled Manufacturing PMI
ggplot(bitcoin_data, aes(x = Date)) +
  geom_line(aes(y = Close, color = "Bitcoin Price")) +
  geom_line(aes(y = manufacturing_pmi_scaled, color = "Manufacturing PMI (scaled)")) +
  ggtitle("Bitcoin Price and Manufacturing PMI Over Time (2020-2024)") +
  xlab("Date") +
  ylab("Value") +
  scale_color_manual(values = c("Bitcoin Price" = "black", "Manufacturing PMI (scaled)" = "blue"))


```

## Address research question

#Initial Reactions to EDA

Based on the initial exploratory data analysis (EDA), it is observed that among the economic indicators analyzed, the Manufacturing PMI (Purchasing Managers' Index) shows the most significant correlation with Bitcoin prices. This correlation is evidenced by the R-squared value derived from the linear regression model, which indicates a substantial relationship between these variables.

#Intuition and Observations

<Correlation Between PMI and Bitcoin Prices>:

The Manufacturing PMI demonstrates the highest R-squared value when compared with Bitcoin prices, suggesting that changes in the PMI are closely related to fluctuations in Bitcoin prices. This observation is crucial as it provides a quantitative measure of the relationship, reinforcing the notion that PMI can serve as a significant economic indicator for predicting Bitcoin price movements.

<Lagging Nature of PMI Trends>:

From the visual representation and analysis, it appears that PMI trends tend to lag behind Bitcoin prices. This lag can be interpreted as PMI reflecting broader economic conditions and business cycles, which eventually influence the liquidity in the market. As business cycles peak and liquidity flows increase, equity-like assets such as Bitcoin tend to reach their peak prices. This lagging nature suggests that PMI could be used as a predictive tool for anticipating future Bitcoin price movements, albeit with a slight delay. Economic Indicators as Business Cycle Representatives:

PMI, as an indicator, represents the overall health of the manufacturing sector and, by extension, the business cycle. During periods of economic expansion, higher PMI values are observed, indicating increased business activity and higher liquidity in the market. This increased liquidity often flows into various asset classes, including equities and cryptocurrencies like Bitcoin. Therefore, a high PMI is often associated with bullish trends in Bitcoin prices, as observed in the initial EDA.

#Research Questions

Can we develop a more robust predictive model that incorporates multiple economic indicators to forecast future prices of Bitcoin? How does the inclusion of these indicators improve the accuracy of price predictions?


